---
title: "R - Assignment Mod 2 Week 2"
author: "Michael Womble"
date: "1/27/2022"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=F}
# The package "ezids" (EZ Intro to Data Science) includes a lot of the helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
# Once installed, load the library.
library(ezids)
```


```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, results = "markup", message = F)
#knitr::opts_chunk$set(warning = F, results = "hide", message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r}
# 1. Answer each question using in words/paragraph. 
# 2. DO NOT use comments inside code blocks (like here) to answer anything. Those are for notes between coders/self. They 
# will be ignored, and not counted as answers. 
# 3. Keep the line/option    knitr::opts_chunk$set(warning = F, results = "hide", message = F)
# You can temporarily turn this on/off and use other option while you are working on the homework if it helps. 
# The submitted work should have this option selected instead.
# 4. All charts/graphs/tables should have appropriate captions. 
```

# HW - Module 2 week 02

## Q1. 
Read in the file with all the baseball players height, weight and age info (BaseballHeightWeight.csv) as a dataframe and save it as baseball. Then display the structure of the dataframe.

```{r Q1, include=T}
baseball<-data.frame(read.csv("BaseballHeightWeight.csv"))
str(baseball)
```
The baseball data is shown to be a data frame with 1034 players listed (observables) and listing 6 variables for each player.

## Q2. 
print out the first 5 and the last 3 rows of the dataframe
```{r Q2}
head(baseball, 5)
tail(baseball, 3)
```

## Q3. 
Change the headers/column names from Weight.pounds. to weight, from Height.inches. to height, and Age to age.
```{r Q3}
colnames(baseball)[which(names(baseball)=="Weight.pounds.")]<-"weight"
colnames(baseball)[which(names(baseball)=="Height.inches.")]<-"height"
colnames(baseball)[which(names(baseball)=="Age")]<-"age"
```

## Q4. 
Print out the first two lines again to see the new column names.
```{r Q4}
head(baseball, 2)
```

## Q5. 
Find the statistics (summary) of the heights, weights, and ages of the players. Using complete sentences, give the reader these summary.
```{r Q5}
summary(baseball)
```
The distribution of player heights has a mean equal to 73.7 inches, a median equal to 74 inches. The tallest player is 83 inches tall and the shortest player is 67 inches.

The distribution of player weights has a mean equal to 202 pounds, a median equal to 200 pounds. The heaviest player is 290 pounds and the lightest player is 150 pounds.

The distribution of player ages has a mean equal to 28.7 years, a median equal to 27.9 years. The oldest player is 48.5 years old and the youngest player is 20.9 years old.

## Q6. 
You should find an issue with the weight data with an NA value. Use the subset( ) function, select that data point (save as shyboy) and find out who the shy boy was.
```{r Q6}
shyboy<-subset(baseball,is.na(weight))
shyboy
```
The shyboy is Kirk Saarloos.

## Q7. 
Remove shyboy from baseball, and then show the number of observations of the cleaned up dataframe baseball.
```{r, Q7}
baseball_clean=baseball[-c(641)]
str(baseball_clean)
```

## Q8. 
Make a quick boxplot of the weight of the players. Use the title( ) function immediately to add a title to the chart. (Try color #0000FF for blue.)
```{r Q8}
boxplot(baseball_clean$weight, col="Blue")
title("Boxplot of Baseball Player Weights")
```

## Q9. 
Make a histogram of the weight of the players. Try color 'green' this time. The function understands the word 'green' just fine. Give it a better title than the default. Does the distribution looks normal?
```{r Q9}
hist(baseball_clean$weight, main='Histogram of Baseball Player Weights', col='Green')
```

## Q10. 
Use ggplot, make a plot with weights vs heights of the players, color by Teams. Give it an appropriate title.
```{r Q10}
library(ggplot2)
ggplot(data=baseball_clean, mapping=aes(x=height, y=weight))+geom_point(aes(color=Team))+labs(title='Player Weight vs Height by Team')
```

## Q11. 
Make a plot with weights vs heights of the players, color by age
```{r Q11}
ggplot(data=baseball_clean, mapping=aes(x=height, y=weight))+geom_point(aes(color=age))+labs(title='Player Weight vs Height by Team')
```

## Q12. 
Obtain a subset of the data with only Team Washington. Using complete sentences, give the number of players in Team Washington.
```{r Q12}
team_washington<-subset(baseball_clean, Team=='WAS')
str(team_washington)
```
There are 36 players on Washington

## Q13. 
Obtain another subset with Team Washington and only players older than 28. How many are there in this veteran group?
```{r Q13}
veteran_washington<-subset(team_washington, age>28)
str(veteran_washington)
```
There are 14 players over the age of 28 on Washington.