---
title: "Intro to DS - PCA / PCR"
author: "Michael Womble"
date: "today"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
library(dplyr)
library(corrplot)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "markup", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```


# USDA Nutrient Dataset  
This data was from USDA National Nutrient dataset, pre-processed by @craigkelly from data.world. 

## Preparation  

### Question 1  
**Import the dataset**  
Import the dataset into R. There should be 45 variables. Take a quick look at the data. Just str, head, and tail will do for now.
```{r impor dataset}
nutrient<-data.frame(read.csv('USDANationalNutrient.csv'))
str(nutrient)
xkabledplyhead(nutrient, title = 'Nutrient Data Frame Head')
xkabledplytail(nutrient, title = 'Nutrient Data Frame Tail')
```
**Answer:** Data is imported as "nutrient". The data frame structure, head, and tails are shown above. All variables, except the first seven variables, are numerical.

### Question 2  
**Drop the first 7 variables**  
Save the rest as a dataframe, and call it `nndf1`. 
```{r subset nndf1}
nndf1<-nutrient[,-c(1:7)]
str(nndf1)
```
**Answer:** First seven variables in the nutrient data frame were dropped and the new data frame is named "nndf1". All remaining variables of numerical data type.

## OLS full model  

### Question 3  
**Issue with full OLS models**  
All the variables are now numerical. Let us build a quick model (full model) to predict `Energy_kcal`, and another full model to predict `VitB12_mcg`. (No fancy output is required.) Vegetarian diet can often be low on B12, which makes one get tired easier. Having a model could be useful. What are the NAs for the coefficients mean? (Try google.) 
```{r full ols model}
ekcal_lm<-lm(formula=nndf1)
ekcal_lm

vitb12df<- nndf1 %>% relocate(VitB12_mcg, .before = Energy_kcal) 
vitb12_lm<-lm(formula=vitb12df)
vitb12_lm
```
**Answer:** The two model outputs are shown above. The NA values in the model output tell us that those variables are linearly related to one or some of the other variables in the data frame.

### Question 4  
**Fix duplicate/collinear variables**  
You probably realize where the problem comes from. Alright then, let us  

* only keep the first 23 columns, 
* standardize all the variables, and 
* the save the resulting dataframe as `nndf`  

Obtain a simple correlation matrix for the dataframe. If you use the function `xkabledply()` on the correlation matrix, it will print out a nice scrollable table for html and your console output.  
```{r collinear variables}
nndf<-nndf1[, c(0:23)]

nndf_corr<-cor(nndf)
xkabledply(nndf_corr, title = 'nndf Correlation Matrix', bso = 'responsive')
```
**Answer:** Since the last 15 variables in the "nndf1" data frame are linearly related to the first 23 variables only the first 23 variables are kept in the data frame "nndf". Above the correlation matrix for the 23 variables is shown.

### Question 5  
**VitB12_mcg**  
Build a quick full model again for “VitB12_mcg”. What is the $R^2$ value? Notice that using PCA won’t improve the $R^2$. It will only match it if you use all 22 PC-components (1 target, 22 explanatory variables). However, this can be the baseline number to compare.  
```{r cor vitB12}
vitb12_df<-nndf %>% relocate(VitB12_mcg, .before = Energy_kcal)
vitb12_lm<-lm(formula = vitb12_df)
summary(vitb12_lm)
```
**Answer:** Looking at the summary of the linear model of vitb12, the model has an R squared value of 0.587.

## PCA  

### Question 6   
**PCA decomposition**  
Instead of PCR, let us run the PCA on this dataset. The dataframe is already centered and scaled. Use  
`nnpcomp <- prcomp(nndf)` to get the 23 PC-components. For `PC1`, which are the four strongest components it consists of?  
(Ans: Riboflavin_mg at -0.3413, Niacin_mg at -0.3378, and so forth.)  
And which are the four strongest for `PC2`? 
```{r PCA decomposition}
nnpcomp<-prcomp(nndf, scale.=TRUE)
nnpcomp
```
**Answer:** The four strongest components of PC1 are Riboflavin_mg, Niacin_mg, VitB6_mg, and Iron_mg with values of -0.3413, -0.3378, -0.3157, -0.2999, respectively. The four strongest components of PC2 are Carb_g, Sugar_g, VitB12_mcg, Protein_g with values of 0.4434, 0.3558, -0.3550, -0.3434, respectively.

### Question 7  
**BiPlots**  
Make a biplot between `PC1` and `PC2` like this: `biplot( nnpcomp, c(1,2), scale=0 )`. Continue to plot `PC1` with `PC3` (use `c(1,3)`), `c(1,4)`, `c(1,5)`, and `c(1,6)`. Also make biplots for `c(2,3)`, `c(3,4)`, `c(4,5)`, `c(5,6)`.
```{r BiPlots, fig.show='hold', out.width='33%'}
nnpcomp12<-biplot(nnpcomp, c(1,2), scale=0)
nnpcomp13<-biplot(nnpcomp, c(1,3), scale=0)
nnpcomp14<-biplot(nnpcomp, c(1,4), scale=0)
nnpcomp15<-biplot(nnpcomp, c(1,5), scale=0)
nnpcomp16<-biplot(nnpcomp, c(1,6), scale=0)
nnpcomp23<-biplot(nnpcomp, c(2,3), scale=0)
nnpcomp34<-biplot(nnpcomp, c(3,4), scale=0)
nnpcomp45<-biplot(nnpcomp, c(4,5), scale=0)
nnpcomp56<-biplot(nnpcomp, c(5,6), scale=0)
```
**Answer:** Above are the biplots.

### Question 8  
**PVE**  
Does the plots show that `PC1` capture much more variance than `PC2`? `PC6`? (Of course we can confirm these with the PVE plot and calculations if we like. It is not required in this hw here.)  
```{r PVE}
nnpcomp_var<-(nnpcomp$sdev^2)
nnpcomp_pve<-nnpcomp_var/sum(nnpcomp_var)
plot(cumsum(nnpcomp_pve), xlab='Principle Component (standardized)', ylab=' Cumulative Proportion of Variance Explained', ylim=c(0,1), type='b')
```
**Answer:** Most of the variance is captured by PC1, about 0.8, while adding PC2 adds a proportion of variance value of 0.1. Combined PC1 and PC2 capture 90% of the variance. Going all the way up to PC6 captures almost all of the variance.

### Question 9  
**Interesting plots**  
If you look at the plot between `PC1` and `PC2`, it looks like a boomerang shape. (See image file in the hw zip.) Recall that `PC1` and `PC2` consists of rather different linear combinations of the original variables. And `PC1`-`PC2` have zero correlation. The chart shows us although `PC1` and `PC2` are overall un-correlated, when `PC1` change, there are only two different trends that `PC2` will change, as if there are only two types of nutrients in the 8000+ observations. Maybe these two kinds of food/items should be analyzed separately with two different models. These are some of the patterns and structures that PCA might be able to reveal.  

At any rate, do you see other charts showing unexpected patterns?

**Answer:** The result discussed above in the question 9 preface is also shown in the plots between PC1 and PC2 - PC6. None of the components, PC2 - PC6 are correlated with PC1, yet when PC1 changes only two trends, at most, the PC2 - PC6 will change.
     
## Optional  

We will skip the model building using PCR here as it turns out not a productive exercise in that direction. While we are having this huge dataset, with 23 variables, this is actually a good candidate for performing the feature selection using leaps.  

If you just look at the dataset and guess what are the good predictors to model Energy_kcal, what would you pick?  

Now try run it by leaps regsubsets. What is the minimum number of predictors you will need, if you want to have a 90% $R^2$?  


```{r, include=FALSE}
# unloadPkg(???)
```



