---
title: "Intro to DS - Logit Regression"
author: "Michael Womble"
date: ""
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

# HW Assignment - Logit Regression

We have the historic Titanic dataset to study here. The version presented has these variables: 

* `survival`: Survival,	0 = No, 1 = Yes
* `pclass`: Ticket class, 1 = 1st, 2 = 2nd, 3 = 3rd
* `sex`: Gender / Sex
* `age`: Age in years
* `sibsp`: # of siblings / spouses on the Titanic
* `parch`: # of parents / children on the Titanic
* `ticket`: Ticket number (for superstitious ones)
* `fare`: Passenger fare
* `embarked`: Port of Embarkment	C: Cherbourg, Q: Queenstown, S: Southampton

The questions listed here are the basic guidelines, not the only goal, in this homework. For example, after you load the dataframe, even though the question does not ask you to look at the structure of the dataframe, you most likely should. You are given less and less “specific to-dos” in the homework, as you are getting more familiar with the data analytic process. Calculate and figure out the necessary info needed in the analysis, even though the questions might not ask for them explicitly. When you look at your own work, you should find it convincing, answering the questions and technically sound.  


## Titanic Tragedy Dataset  

### Question 1

**Import the dataset into R**  
Import the dataset into R, and call it titanic_orig, and explore it little to get the overall picture of the dataset. Eventually we would like to see what affected survival in the tragedy. 
```{r importing titanic dat}
titanic_org<-data.frame(read.csv('Titanic.csv'))
```
**Answer:** Data set was imported as a data frame with the name titanic_orig.

### Question 2 
**Age**  
One of the main factors we will try is Age. How many missing values are there for the variable `age`? If not too many, we should just clean those up and subset those out.
```{r removing missing values from age variable}
summary(titanic_org)
str(titanic_org)
age_missing<-subset(titanic_org, is.na(age))
titanic_org<-na.omit(titanic_org)
summary(titanic_org)
```
**Answer:** The age variable had 177 missing values. These were subsetted out and removed fromt the dataset.


### Question 3  
**More clean up**  
While we are cleaning up the data, if we were to use sibsp and parch in our analysis, even though they are legitimately ratio level variables, we might not expect doubling the number of siblings necessarily double the effects on survival. For this reason, we should change these two variables to factor levels. Also change the other ones that you find imported as the wrong data type.  
```{r clean up}
str(titanic_org)
titanic_org$sibsp<-as.factor(titanic_org$sibsp)
titanic_org$parch<-as.factor(titanic_org$parch)
titanic_org$sex<-as.factor(titanic_org$sex)
titanic_org$embarked<-as.factor(titanic_org$embarked)
str(titanic_org)
```
**Answer:** I changed four variables to the fator data type. These variables are the sibsp, parch, sex, and embarked variables.


## Pre-logistic Regression

### Question 4  
**Survival and age**  
Before using our newly learned technique with logistic regression, let’s go old school, and use some prior knowledge to try find an answer. Does the data support that `age` very much affects `survival`?
```{r age vs survival, results='markup'}
age_survival_table<-xtabs(~ survived + age, data = titanic_org)
age_survival_table

chisqr_age<-chisq.test(age_survival_table)
chisqr_age
```

```{r age vs survival boxplot}
library(ggplot2)

boxplot(titanic_org$age~titanic_org$survived, 
        main='Age of Titanic Passengers',
        xlab = 'Survival',
        ylab = 'Age',
        col=c('Red','Green'))

age_survival<-subset(titanic_org, survived==1)
age_dead<-subset(titanic_org, survived==0)
t.test(age_survival$age, mu=mean(age_dead$age), conf.level = 0.95)
```
**Answer:** To test for variable dependency a contingency table comparing survival and age was made. The chi square test was then run on the table giving a p-value of 0.1 indicating the variables are independent of each other. However, the expected cell values for many of the cells was less than 5 meaning that this may be unreliable. To double check a boxplot was made comparing the age distributions of the passengers that survived and those that did not. There appears to be no difference between the two. To further confirm this the data was subsetted to form a survivor group and a dead group. A t-test comparing the age mean of the two subsets. This resulted in a p-value of 0.01 indicating that there is a difference between the means of these two groups which means there may be a dependency of survival on age.

### Question 5  
**Survival and gender**  
Similarly, does the data support `sex` has an effect on `survival`? 
```{r sex vs survival, results='markup'}
sex_survival_table<-xtabs( ~ survived + sex, data = titanic_org)
sex_survival_table

chisqr_sex<-chisq.test(sex_survival_table)
chisqr_sex
```
**Answer:** Making a contingency table and running a chi square test on it results in a p-value of < 2E-16 indicating a dependency of survival on sex.

### Question 6   
**Survival and pclass**  
Another big question is, does the data support Ticket class `pclass` has an effect on `survival`? 
```{r pclass vs survival, results='markup'}
pclass_survival_table<-xtabs( ~ survived + pclass, data=titanic_org)
pclass_survival_table

chisqr_pclass<-chisq.test(pclass_survival_table)
chisqr_pclass
```

**Answer:** Making a contingency table and running a chi square test on it results in a p-value of < 2E-16 indicating a dependency of survival on passenger class.


## Logistic Regression

### Question 7   
**Survival and age + pclass**  
Now let us build a logit model with age+pclass as predictors, and analyze the results. Remember to do all the model evaluation steps. Is the model a good one?
```{r logistic regress model 1, results='markup'}
loadPkg("regclass")
loadPkg("ResourceSelection")
loadPkg("pROC")

survival_logit_model1<-glm(survived ~ age + pclass, data=titanic_org, family = 'binomial')
summary(survival_logit_model1)

prob1=predict(survival_logit_model1, type = "response" )
titanic_org$prob1=prob1
h1 <- roc(survived~prob1, data=titanic_org)
auc(h1) # area-under-curve prefer 0.8 or higher.
plot(h1)
```
**Answer:** This model is a good one. Measures of this are the AIC value and the area under the curve value, for this model those values are 833.4 and 0.744, respectively. The ideal value for area under the curves is ~ 0.8. With our value of 0.744 we are already pretty close to the optimum value.

### Question 8  
**More features**  
Can we improve the model? Let us also throw in `sex` as a predictor. How’s the model now?
```{r logit regression model 2, results='markup'}
survival_logit_model2<-glm(survived ~ age + pclass+ sex, data=titanic_org, family='binomial')
summary(survival_logit_model2)

prob2=predict(survival_logit_model2, type = "response" )
titanic_org$prob1=prob2
h2 <- roc(survived~prob2, data=titanic_org)
auc(h2) # area-under-curve prefer 0.8 or higher.
plot(h2)
```
**Answer:** With the addition of the sex predictor our model improves further still. With AIC and area under the curve values of 655.3 and 0.852, respectively. Model 2 has a reduced AIC value compared to model 1 indicating a better fit. The area under the curve also increases from 0.744 to 0.852 in model 2, also indicating a better fitting model. However, with an area under the curve value of 0.852, model 2 is on the side of overfitting compared with the optimum value of 0.8, although it is marginally closer to 0.8, than model 1 is.

### Question 9  
**Sample Predictions**  
According to the last model, what is the chance of survival for a female, age 10, second class passenger? And a male, age 20, first class passenger?
```{r applying model}
female_10<-data.frame(age=10, pclass=2, sex='female')
male_20<-data.frame(age=20, pclass=1, sex='male')

predict(survival_logit_model2, newdata = female_10, type='response')
predict(survival_logit_model2, newdata = male_20, type='response')
```
**Answer:** According to model 2 the 2nd class 10 year old female has an 89.5% chance of surviving while the 1st class 20 year old male has a 62.4% chance of surviving.

## Interpretation  

### Question 10  
*Summary*  
With all the results you obtained above, how would you present a high-level summary of the findings? Are the results surprising or expected? You might need to dig a little deeper than just the numbers, the test results/p-values, and the model statistics. This is a question about who we are, … in the face of death. 


```{r, include=FALSE}
# unloadPkg(???)
```
**Answer:**


